\myTitle

### 2014-08-21 tex/abstract.tex
Fixed some mathematical sections then submitted.
On a last reading I found some bone-headed errors in
the mathematical section --- I left the \ell subscript
on things when it shouldn't have been.  So, I cleaned
that up and fixed it.

Then included referees in the cover letter and fixed a
few typos and submitted it.

This is the version that was submitted.

      We show that the the algorithm in the program {\sc flock} \citep{Duc&Tur2009} can be
interpreted as an estimation procedure using  
a model that is essentially identical to the {\sc structure} 
\citep{Pritchardetal2000} model with no admixture and non-correlated 
allele frequency priors. The computational routine in {\sc flock} is 
equivalent to searching for the maximum-{\em a-posteriori} (MAP) 
estimate of this restricted {\sc structure} model via a simulated 
annealing algorithm with an extreme cooling 
schedule (namely, the exponent on the objective function~$\rightarrow \infty$).  By analysis of a large, real data set we
confirm that {\sc flock} yields results similar to {\sc structure} with the no-admxixture
model and uncorrelated allele frequency prior with more variability 
from run to run. Interpreting {\sc flock}'s algorithm as a special case of the model in 
{\sc structure} should aid in  understanding the program's output and behavior.

>>>>

      We show that the the algorithm in the program {\sc flock} \citep{Duc&Tur2009} can be
interpreted as an estimation procedure using  
a model that is essentially identical to the {\sc structure} 
\citep{Pritchardetal2000} model with no admixture and non-correlated 
allele frequency priors. The computational routine in {\sc flock} is 
equivalent to searching for the maximum-{\em a-posteriori}
estimate of this restricted {\sc structure} model via a simulated 
annealing algorithm with an extreme cooling 
schedule (namely, the exponent on the objective function~$\rightarrow \infty$).  By analysis of a large, real data set we
confirm that {\sc flock} yields results similar to {\sc structure} with the no-admxixture
model and uncorrelated allele frequency prior with more variability 
from run to run. Interpreting {\sc flock}'s algorithm as a special case of the model in 
{\sc structure} should aid in  understanding the program's output and behavior.

### 2014-08-21 tex/flock-comment-letter-to-editor.tex
Fixed some mathematical sections then submitted.
On a last reading I found some bone-headed errors in
the mathematical section --- I left the \ell subscript
on things when it shouldn't have been.  So, I cleaned
that up and fixed it.

Then included referees in the cover letter and fixed a
few typos and submitted it.

This is the version that was submitted.

In two publications {\sc flock}'s authors have been somewhat outspoken in asserting that {\sc flock} is very different from {\sc structure}. Indeed, we embarked upon this project because numerous colleagues were recommending we try the program {\sc flock} in cases where {\sc structure} did not reveal much population structure.  Upon a close reading of the mathematical details of {\sc flock}'s algorithm, however, it was clear that {\sc flock} is very similar to {\sc structure}.  We describe this from a mathematical perspective and then perform analyses on a real data set to confirm it.  These analyses additionally expose some very real problems suffered by the {\sc flock} approach to estimating $K$, the number of clusters.  While we feel our findings would easily substantiate a position that there were few, if any, reasons to use {\sc flock} instead of {\sc structure}, we don't assert that point strongly in the paper. Rather, in consideration of the work that has been done to develop {\sc flock}, we try to adopt a neutral tone, stating our mathematical and simulation findings as plainly as possible and allowing readers to make up their own minds.  Of course, if the editors felt that a stronger stance were appropriate, we could provide that, as well.

@@

>>>>

In two publications {\sc flock}'s authors have been somewhat outspoken in asserting that {\sc flock} is very different from {\sc structure}. Indeed, we embarked upon this project because numerous colleagues were recommending we try the program {\sc flock} in cases where {\sc structure} did not reveal much population structure.  Upon a close reading of the mathematical details of {\sc flock}'s algorithm, however, it was clear that {\sc flock} is very similar to {\sc structure}.  We describe this from a mathematical perspective and then perform analyses on a real data set to confirm it.  These analyses additionally expose some real problems that make it difficult to apply the {\sc flock} approach to estimating $K$, the number of clusters.  While we feel our findings would easily substantiate a position that there were few, if any, reasons to use {\sc flock} instead of {\sc structure}, we don't assert that point strongly in the paper. Rather, in consideration of the work that has been done to develop {\sc flock} and in the spirit of us all being molecular ecologists working together to understand the tools we use, we try to adopt a neutral tone, stating our mathematical and simulation findings as plainly as possible and leaving it to the reader to make up his or her own mind about which programs to use.  Of course, if the editors felt that a stronger stance were appropriate, we could provide that in a revised draft.

@@

We have given some thought to who would be good referees for this paper.  Given the technical and terse nature of our comparison between {\sc flock} and {\sc structure}, it would be advantageous to enlist referees who are very familiar with the model underlying {\sc structure}.  Obvious choices would be Jonathan Pritchard, Matthew Stephens, and John Novembre.  However, since these are very busy investigators running large human statistical-genetic labs, other possibilities would be Melissa Hubisz, Bruce Rannala, Greg Wilson (author of BayesAss), Philip Andolfatto, John Hulsenbeck, Jerome Pella, or Michelle Masuda.  Specifically in the molecular ecology realm, we think Oscar Gaggiotti would make a very good referee.

Daniel Falush would be the most appropriate editor for this manuscript because he has contributed to the development of {\sc structure} and is extremely familiar with the underlying model.

### 2014-08-21 tex/main-body-text.tex
Fixed some mathematical sections then submitted.
On a last reading I found some bone-headed errors in
the mathematical section --- I left the \ell subscript
on things when it shouldn't have been.  So, I cleaned
that up and fixed it.

Then included referees in the cover letter and fixed a
few typos and submitted it.

This is the version that was submitted.

\section*{Introduction}
Methods for the unsupervised clustering of genotypes have received 
considerable attention in the molecular ecology literature.  
The best known example of this class of methods is {\sc structure} 
\citep{Pritchardetal2000}, with over 12,000 citations identified on 
Google Scholar.  A variety of other 
clustering methods have been developed that are all closely related to 
{\sc structure}, for example {\sc NewHybrids} \citep{And&Tho2002}, {\sc 
BayesAss+} \citep{Wil&Ran2003}, and {\sc baps} 
\citep{Coranderetal2004}. An overview of the similarities between these methods can be 
found in \citet{Anderson2009PGAC}.

@@

However, here we identify 
the close relationship between {\sc flock}'s algorithm and the program
{\sc structure}.  Although the algorithm in {\sc flock}
is, as the authors point out, deterministic (apart from a random initial allocation of 
individuals to cluster), it can be interpreted as a 
limiting case of a probabilistic search algorithm applied to one of {\sc structure}'s
models.  More specifically, the {\sc flock} 
algorithm is a special case of the simulated annealing
algorithm for finding the Bayesian maximum-{\em a-posteriori}
estimate from a marginalized form of the {\sc structure} model with no
admixture and non-correlated allele frequencies.

@@

It will be useful for our comparison with {\sc flock} to point out that another way 
of pursuing inference in this model would be to first integrate out the 
Dirichlet priors on the allele frequencies and then sample from the posterior
for the $Z_i$'s by Gibbs sampling.  When $\theta$ is integrated out, the genotypes 
of the individuals are no longer conditionally independent (they were originally
independent {\em conditional} on $\theta$ and the $Z_i$'s), so the calculation of 
the full conditional distribution of $Z_i$ becomes somewhat more involved,
but can be computed by the following reasoning. First, the 
conditional distribution of $Y_{i\ell}$, given that individual $i$ is from subpopulation $k$
(\ie $Z_i=k$), depends on the 
cluster memberships and genotypes  of all the remaining individuals,
which we denote by  $Z_{(-i)}$ and $Y_{(-i)\ell}$, respectively.
This full conditional for $Y_{i\ell}$ follows 
a Compound Dirichlet Multinomial distribution (CDM):
\begin{eqnarray}
\lefteqn{(Y_{i\ell}~|~Z_i=k,~Z_{(-i)},~Y_{(-i)\ell}) \sim} \label{eq:cdm}\\
& & \mathrm{CDM}(\lambda_{k\ell 1} + \#_{(-i)}(k,\ell,1), \ldots,
\lambda_{k\ell A_\ell} + \#_{(-i)}(k,\ell,A_\ell)), \nonumber
\end{eqnarray}
where $\#_{(-i)}(k,\ell,j)$ is the number of alleles of type $j$ at locus $\ell$
found in individuals {\em other than individual $i$} that currently
belong to cluster $k$. Thus, calculating (\ref{eq:cdm}) for each value of $Z_i=k \in 
\{1,\ldots,K\}$ 
and normalizing to sum to one gives the full conditional distribution for 
$Z_i$:
\begin{equation}
P(Z_i=k~|~Y_i, ~Z_{(-i)},~Y_{(-i)\ell})~~,~~k=1,\ldots,K.
\label{eq:fc}
\end{equation}
A new value of $Z_i$ would be drawn from (\ref{eq:fc}) if doing Gibbs sampling in this
version of the {\sc structure} model in which $\theta$ has been integrated out.  In 
fact,
this is the approach (with a slightly different prior weight on the $Z_i$'s) taken to 
update 
$Z_i$ in both {\sc hwler} \citep{Pel&Mas2006} and {\sc structurama} \citep{Hue&And2007} 
when not
proposing changes to the number of clusters.

@@

From that description, it is not clear how \citeauthor{Duc&Tur2009} treat 
alleles that appear in the focal
individual but do not appear within any other individuals within a cluster.
The original approach of \citet{Paetkauetal1995} merely added 0.01 or some other
small value to each allele frequency that was 0, and then renormalized 
the allele frequencies to sum to 1.  In later works, Paetkau and colleagues
also employed the ``Bayesian'' approach of \citet{Ran&Mou1997} in their 
software {\sc geneclass2}  \citep{Piryetal2004}.  If \citeauthor{Duc&Tur2009}
use the approach of \citet{Ran&Mou1997}, then the $i\thh$ individual will
be reallocated to whichever cluster gives the highest value to 
$P(Y_{i\ell}~|~Z_i=k,~Z_{(-i)},~Y_{(-i)\ell})$, which is exactly the probability
defined in (\ref{eq:cdm}).  
If \citeauthor{Duc&Tur2009} use the simpler ``add 0.01 and renormalize''
formulation of \citet{Paetkauetal1995}, then their reallocations are based
on maximizing a likelihood that, although it may not be formally identical to
that in (\ref{eq:cdm}), is nearly identical to it.

It is worth pointing out that, since the
{\sc structure} model without admixture assumes a uniform prior over 
the $K$ different clusters, $P(Y_{i\ell}~|~Z_i=k,~Z_{(-i)},~Y_{(-i)\ell})$
is exactly proportional to $P(Z_i=k~|~Y_i, ~Z_{(-i)},~Y_{(-i)\ell})$, which 
we have shown is the
full conditional distribution for $Z_i$ in the {\sc structure} model after
integrating out $\theta$.  Thus, we have shown that, doing Gibbs
sampling for $Z_i$ in  the {\sc structure} model in which $\theta$ has been
integrated out would involve sampling a new value of $Z_i$ from the 
probability distribution,
\[
P(Z_i=k~|~Y_i, ~Z_{(-i)},~Y_{(-i)\ell})
\]
and that the updates that
{\sc flock} makes to each $Z_i$ involve assigning to $Z_i$
\begin{equation}
\arg\max_k P(Z_i=k~|~Y_i, ~Z_{(-i)},~Y_{(-i)\ell}),
\end{equation}
\ie the value of $k$ that maximizes $P(Z_i=k~|~Y_i, ~Z_{(-i)},~Y_{(-i)\ell})$
(or, as stated above, a nearly identical function of $k$, depending on how
the program {\sc flock} treats alleles that are not observed in certain clusters).

\subsection*{The {\sc flock} algorithm and {\sc structure}}
The goal of {\sc structure}'s MCMC algorithm is to sample from the space of 
$Z$'s in proportion to their posterior probability.  If, all that was desired
was a point estimate of the $Z_i$'s that maximized the posterior probability,
then one standard approach to finding that maximum-{\em a-posteriori}, or MAP, 
estimate, would be to use simulated annealing \citep{Kirkpatricketal1983}.
To do simulated annealing in the {\sc structure} model with no admixture and
no correlated allele frequency prior, after integrating out the unknown
allele frequencies, the updates for each $Z_i$ would be made from the 
distribution
\[
\biggl[P(Z_i=k~|~Y_i, ~Z_{(-i)},~Y_{(-i)\ell})\biggr]^\beta
\]
which is the same full conditional distribution as the marginalized {\sc structure}
model (equation~\ref{eq:fc}), raised
to the power $\beta$.  In typical applications of simulated annealing,
$\beta$ is set to a starting value less than 1, and, as the algorithm proceeds,
$\beta$ is increased according to a ``cooling schedule''\citep{Hajek1988}, until
the algorithm finds a local maximum in the posterior probability surface.  
When $\beta$ is small, the process can easily traverse ``valleys''
in the posterior probability surface and has a better chance of searching
more of the space for the maximum value.  As $\beta$ gets larger, it becomes
more probable that the process will
move uphill on the probability surface, tending toward the local
maximum near where it currently is.

@@

{\sc flock} and {\sc structure} provide similar results in 
terms of how individuals are allocated among runs at low values of $K$, 
while at larger values, {\sc flock} often found, in addition to the solution identified by {\sc structure}, 
multiple other solutions. For all six groups 
of runs with $K = 2$ to 4, {\sc flock} obtained a nearly identical clustering of individuals 
(See supplementary materials) none of which differed substantially from {\sc structure}'s results
(Fig.~\ref{fig:qplots}). Of the `best runs' for the six groups for $K$ = 2, two  
runs yielded identical clustering (same mean LLOD score) while all other runs 
differed by at most the allocation of 13 individuals. As 
$K$ increased so did the number of solutions
that {\sc flock} found (Fig 1). For all groups of runs for $K > 2$, no plateaus were observed---all intial 
allocations resulted in solutions that differed. 
{\sc flock} still produced estimates of the LLOD by individual for the 
'best run', however, it is unclear in these cases what constitutes the 'best run' as all plateau lengths are 
equal to one. Many of these solutions are, however, surprisingly similar to
the partitions identified by {\sc structure}. Both programs idenfity clusters 
which can be attributed to the geographical boundaries identified by \citet{Garzaetal_norcal},
although less reliably so in the case of {\sc flock}.

@@

Though we observed similar results
between the two programs at low values of $K$, {\sc flock}'s `best-run' solutions became somewhat inconsistent at higher values of $K$.
This observation was consistent with \citet{Duc&Tur2012}'s comparison of the 
two programs using simulated data 
for $K=4$ \& 8, $L=10$ (Fig. 4 in \citet{Duc&Tur2012}). 
This is likely a consequence of the rapidly increasing size of the search space over possible 
partitions as $K$ and $n$, the number of individuals, increase. 
The likelihood surface defined on partitions of this space will often
be multimodal---possessing numerous local peaks and troughs.
The peak that  {\sc flock}'s algorithm finds is entirely determined by the initial, random, allocation of 
individuals, as it is a deterministic algorithm.
If the number of initial allocations does not increase
as $K$ is increased, the starting allocations become diffusely scattered in the space of all
possible partitions making it less likely that many (if any at all) will converge to the same local mode. 
This coupled with the fact that there is 
ambiguity in what defines a `best run' when all plateaus are of length zero complicates the interpretation
of results from {\sc flock} if no non-zero plateaus are encountered. The problem of getting 
caught in local modes is not unique to {\sc flock}---{\sc structure} also can converge to 
different solutions in complex spaces; however our results suggest this is a bigger problem for {\sc flock}.
By increasing the number of loci used in their simulations, \citet{Duc&Tur2012} showed  that 
{\sc flock} was able to converge in a simulated data set with $K=8$, but held no advantage in inference of $K$ over {\sc structure}. That simulated data set, with $n=240$ was considerably smaller and less complex than the real data investigated here.
%They simulated data using 500 individuals per population but only used 30 of them in the analysis.
% if you have more loci than individuals per population with high Fst between groups it does great!

>>>>

\section*{Introduction}
Methods for the unsupervised clustering of genotypes have received 
considerable attention in the molecular ecology literature.  
The best known example of this class of methods is {\sc structure} 
\citep{Pritchardetal2000}, with over 12,000 citations identified on 
Google Scholar.  A variety of other 
clustering methods have been developed that are all closely related to 
{\sc structure}, for example {\sc NewHybrids} \citep{And&Tho2002}, {\sc 
BayesAss+} \citep{Wil&Ran2003}, and {\sc baps} 
\citep{Coranderetal2004}. An overview of the similarities between several of these methods can be 
found in \citet{Anderson2009PGAC}.

@@

However, here we identify 
the close relationship between {\sc flock}'s algorithm and the program
{\sc structure}.  Although the algorithm in {\sc flock}
is, as the authors point out, deterministic (apart from a random initial allocation of 
individuals to clusters), it can be interpreted as a 
limiting case of a probabilistic search algorithm applied to one of {\sc structure}'s
models.  More specifically, the {\sc flock} 
algorithm is a special case of the simulated annealing
algorithm for finding the Bayesian maximum-{\em a-posteriori}
estimate from a marginalized form of the {\sc structure} model with no
admixture and non-correlated allele frequencies.

@@

It will be useful for our comparison with {\sc flock} to point out that another way 
of pursuing inference in this model would be to first integrate out the 
Dirichlet priors on the allele frequencies and then sample from the posterior
for the $Z_i$'s by Gibbs sampling.  When $\theta$ is integrated out, the genotypes 
of the individuals are no longer conditionally independent (they were originally
independent {\em conditional} on $\theta$ and the $Z_i$'s), so the calculation of 
the full conditional distribution of $Z_i$ becomes somewhat more involved,
but can be computed by the following reasoning. First, the 
conditional distribution of $Y_{i\ell}$, given that individual $i$ is from subpopulation $k$
(\ie $Z_i=k$), depends on the 
cluster memberships and genotypes at locus $\ell$ of all the remaining individuals,
which we denote by  $Z_{(-i)}$ and $Y_{(-i)\ell}$, respectively.
This full conditional for $Y_{i\ell}$ follows 
a Compound Dirichlet Multinomial distribution (CDM):
\begin{eqnarray}
\lefteqn{(Y_{i\ell}~|~Z_i=k,~Z_{(-i)},~Y_{(-i)\ell}) \sim} \label{eq:cdm}\\
& & \mathrm{CDM}(\lambda_{k\ell 1} + \#_{(-i)}(k,\ell,1), \ldots,
\lambda_{k\ell A_\ell} + \#_{(-i)}(k,\ell,A_\ell)), \nonumber
\end{eqnarray}
where $\#_{(-i)}(k,\ell,j)$ is the number of alleles of type $j$ at locus $\ell$
found in individuals {\em other than individual $i$} that currently
belong to cluster $k$. Thus, calculating the product of (\ref{eq:cdm}) across all the loci
($\ell = 1,\ldots,L$) gives the full conditional multilocus genotype probability 
for individual $i$:
\begin{eqnarray}
\lefteqn{P(Y_{i}~|~Z_i=k,~Z_{(-i)},~Y_{(-i)}) =}  \label{eq:cdm-multi-prob} \\
& &  \prod_{\ell=1}^L P(Y_{i\ell}~|~Z_i=k,~Z_{(-i)},~Y_{(-i)\ell}). \nonumber
\end{eqnarray}
Computing (\ref{eq:cdm-multi-prob}) for 
for each value of $Z_i=k \in 
\{1,\ldots,K\}$ 
and normalizing to sum to one gives the full conditional distribution for 
$Z_i$:
\begin{equation}
P(Z_i=k~|~Y_i, ~Z_{(-i)},~Y_{(-i)})~~,~~k=1,\ldots,K.
\label{eq:fc}
\end{equation}
A new value of $Z_i$ would be drawn from (\ref{eq:fc}) if doing Gibbs sampling in this
version of the {\sc structure} model in which $\theta$ has been integrated out.  In 
fact,
this is the approach (with a slightly different prior weight on the $Z_i$'s) taken to 
update 
$Z_i$ in both {\sc hwler} \citep{Pel&Mas2006} and {\sc structurama} \citep{Hue&And2007} 
when not
proposing changes to the number of clusters.

@@

From that description, it is not clear how \citeauthor{Duc&Tur2009} treat 
alleles that appear in the focal
individual but do not appear within any other individuals within a cluster.
The original approach of \citet{Paetkauetal1995} merely added 0.01 or some other
small value to each allele frequency that was 0, and then renormalized 
the allele frequencies to sum to 1.  In later works, Paetkau and colleagues
also employed the ``Bayesian'' approach of \citet{Ran&Mou1997} in their 
software {\sc geneclass2}  \citep{Piryetal2004}.  If \citeauthor{Duc&Tur2009}
use the approach of \citet{Ran&Mou1997}, then the $i\thh$ individual will
be reallocated to whichever cluster gives the highest value to 
$P(Y_{i}~|~Z_i=k,~Z_{(-i)},~Y_{(-i)})$, which is 
exactly the probability
defined in (\ref{eq:cdm-multi-prob}).  
If \citeauthor{Duc&Tur2009} use the simpler ``add 0.01 and renormalize''
formulation of \citet{Paetkauetal1995}, then their reallocations are based
on maximizing a likelihood that, although it may not be formally identical to
the probability defined in (\ref{eq:cdm-multi-prob}), is nearly identical to it.

It is worth pointing out that, since the
{\sc structure} model without admixture assumes a uniform prior over 
the $K$ different clusters, $P(Y_{i}~|~Z_i=k,~Z_{(-i)},~Y_{(-i)})$
is exactly proportional to $P(Z_i=k~|~Y_i, ~Z_{(-i)},~Y_{(-i)})$, which 
we have shown is the
full conditional distribution for $Z_i$ in the {\sc structure} model after
integrating out $\theta$.  Thus, we have shown that, doing Gibbs
sampling for $Z_i$ in  the {\sc structure} model in which $\theta$ has been
integrated out would involve sampling a new value of $Z_i$ from the 
probability distribution,
\[
P(Z_i=k~|~Y_i, ~Z_{(-i)},~Y_{(-i)})
\]
and that the updates that
{\sc flock} makes to each $Z_i$ involve assigning to $Z_i$
\begin{equation}
\arg\max_k P(Z_i=k~|~Y_i, ~Z_{(-i)},~Y_{(-i)}),
\end{equation}
\ie the value of $k$ that maximizes $P(Z_i=k~|~Y_i, ~Z_{(-i)},~Y_{(-i)})$
(or, as stated above, a nearly identical function of $k$, depending on how
the program {\sc flock} treats alleles that are not observed in certain clusters).

\subsection*{The {\sc flock} algorithm and {\sc structure}}
The goal of {\sc structure}'s MCMC algorithm is to sample from the space of 
$Z$'s in proportion to their posterior probability.  If, all that was desired
was a point estimate of the $Z_i$'s that maximized the posterior probability,
then one standard approach to finding that maximum-{\em a-posteriori}, or MAP, 
estimate, would be to use simulated annealing \citep{Kirkpatricketal1983}.
To do simulated annealing in the {\sc structure} model with no admixture and
no correlated allele frequency prior, after integrating out the unknown
allele frequencies, the updates for each $Z_i$ would be made from the 
distribution
\[
\biggl[P(Z_i=k~|~Y_i, ~Z_{(-i)},~Y_{(-i)})\biggr]^\beta
\]
which is the same full conditional distribution as the marginalized {\sc structure}
model (equation~\ref{eq:fc}), raised
to the power $\beta$.  In typical applications of simulated annealing,
$\beta$ is set to a starting value less than 1, and, as the algorithm proceeds,
$\beta$ is increased according to a ``cooling schedule''\citep{Hajek1988}, until
the algorithm finds a local maximum in the posterior probability surface.  
When $\beta$ is small, the process can easily traverse ``valleys''
in the posterior probability surface and has a better chance of searching
more of the space for the maximum value.  As $\beta$ gets larger, it becomes
more probable that the process will
move uphill on the probability surface, tending toward the local
maximum near where it currently is.

@@

{\sc flock} and {\sc structure} provide similar results in 
terms of how individuals are allocated among runs at low values of $K$, 
while at larger values, {\sc flock} often found, in addition to the solution identified by {\sc structure}, 
multiple other solutions. For all six groups 
of runs with $K = 2$ to 4, {\sc flock} obtained a nearly identical clustering of individuals 
(See supplementary materials) none of which differed substantially from {\sc structure}'s results
(Fig.~\ref{fig:qplots}). Of the `best runs' for the six groups for $K$ = 2, two  
runs yielded identical clustering (same mean LLOD score) while all other runs 
differed by at most the allocation of 13 individuals. As 
$K$ increased so did the number of solutions
that {\sc flock} found (Fig 1). For all groups of runs for $K > 2$, no plateaus were observed---all initial 
allocations resulted in solutions that differed. 
{\sc flock} still produced estimates of the LLOD by individual for the 
'best run', however, it is unclear in these cases what constitutes the 'best run' as all plateau lengths are 
equal to one. Many of these solutions are, however, surprisingly similar to
the partitions identified by {\sc structure}. Both programs idenfity clusters 
which can be attributed to the geographical boundaries identified by \citet{Garzaetal_norcal},
although less reliably so in the case of {\sc flock}.

@@

Though we observed similar results
between the two programs at low values of $K$, {\sc flock}'s `best-run' solutions became somewhat inconsistent at higher values of $K$.
This observation was consistent with \citet{Duc&Tur2012}'s comparison of the 
two programs using simulated data 
for $K=4$ \& 8, $L=10$ (Fig. 4 in \citet{Duc&Tur2012}). 
This is likely a consequence of the rapidly increasing size of the search space over possible 
partitions as $K$ and $n$, the number of individuals, increase. 
The likelihood surface defined on partitions of this space will often
be multimodal---possessing numerous local peaks and troughs.
The peak that  {\sc flock}'s algorithm finds is entirely determined by the initial, random, allocation of 
individuals, as the algorithm proceeds deterministically after the initial random allocation.
If the number of initial allocations does not increase
as $K$ is increased, the starting allocations become diffusely scattered in the space of all
possible partitions making it less likely that many (if any at all) will converge to the same local mode. 
This coupled with the fact that there is 
ambiguity in what defines a `best run' when all plateaus are of length zero complicates the interpretation
of results from {\sc flock} if no non-zero plateaus are encountered. The problem of getting 
caught in local modes is not unique to {\sc flock}---{\sc structure} also can converge to 
different solutions in complex spaces; however our results suggest this is a bigger problem for {\sc flock}.
By increasing the number of loci used in their simulations, \citet{Duc&Tur2012} showed  that 
{\sc flock} was able to converge in a simulated data set with $K=8$, but held no advantage in inference of $K$ over {\sc structure}. That simulated data set, with $n=240$ was considerably smaller and less complex than the real data investigated here.
%They simulated data using 500 individuals per population but only used 30 of them in the analysis.
% if you have more loci than individuals per population with high Fst between groups it does great!
